### 2020-01-15 paper/sex_selection_spacing-ver3.tex
Incorporate predicted fertility file and set up for make run
Write discussion of fertility prediction and results

>>>>

Birth spacing has increased substantially over time across all education groups, but 
especially among those women likely to use sex selection.
One of the potential effects of the increased birth spacing is that the total fertility
rate may underestimate cohort fertility, at least temporarily.
Whether tempo effects ``biases'' fertility depends on what the outcome of interest is,
but since my outcome of interest is cohort fertility I will use ``bias''
\citep{Ni-Bhrolchain2011}.

Some tempo effects, such as those that arise from postponement of childbearing, are 
studied extensively in the literature \citep{Bongaarts1999}.
There are, however, to the best of my knowledge, no other cases where there has been
as substantial an increasing in spacing as for India.

To understand how cohort fertility has changed over time and how strong the bias in
TFR is, I compare predicted fertility using both a version of the total fertility
rate and the hazard model results presented above as shown in Table \ref{tab:fertility}.

\input{../tables/fertility.tex}

The TFR measure uses births up to parity four to make the measure comparable to what I 
can calculate for the hazard model, and I calculate this four-parity TFR by urban/rural 
and the education groups used above.
Hence, the TFRs are not directly comparable to those in the NFHS reports, although the
differences should be smaller the higher the education group and the most recent the 
survey provided that cohort fertility has fallen over time.
Otherwise, I follow the procedure used in the NFHS reports and calculate age specific 
fertility rates for five-year age groups using the number of births in months 1 to 36 
before the survey month (births in the survey month are not included) and the number
of women in each age group.%
\footnote{
See \citet{Croft2018} for more detail.
Because NFHS-1 and NFHS-2 did not interview all women about fertility, the number of women 
is based on the household rosters, assuming that never-married women have had no children.
Because of the low number of births to women aged 45 to 49 this age group was not
used for the hazard model estimations and, therefore, is also excluded for the fertility
rate.
}

NFHS-1 was collected after sex selection was introduced and it is, therefore, not 
possible to calculate a fertility rate in exactly the same manner for a period before 
sex selection was widely available.
Instead, I calculate the fertility rates for women who was between 15 and 39 years of
age five years prior to the survey month, again using the number of births three years 
before that.
Given the relatively low number of births to women 40 to 45 years of age, this approach 
provides the best estimate of the fertility rate at the time when sex selection still was 
not widespread.
This fertility rate is shown as ``1987--88'' in the table.

To predict cohort fertility based on the models above, I estimate the probability of 
progression to the next parity for each spell.
I include the spell from marriage to first birth, even those results are not examined 
above because of the problems with capturing the exact timing of marriages for the first 
two rounds of the survey.
The estimated progression probabilities should, however, be less affected by this
problem.
To ensure that the predictions depend solely on sample composition and the estimate, I
use the age at marriage for each woman and predict the likelihood of progressing to
each parity, assuming two years increases in age between each parity.
Hence, for a woman married at age 20, I use age 22 for predicting progression 
to the second birth, 24 for the third, and 26 for the fourth.

The parity progression depends on the sex composition for prior children for
parity two and above.
For example, the likelihood of a second birth is the probability of having a first birth 
multiplied with
sum of the 
probability of having a son as the first child conditional on having a first child 
multiplied with parity progression probability if the first child is a son 
and the 
probability of having a girl as the first child  conditional on having a first child
multiplied with the parity progression probability if the first child is a girl.
$$
prob(2nd) = prob(1st) * (prob(son|1st)*prob(2nd|son) + prob(girl|1st) * prob(2nd|girl)
$$

The survey rounds do not coincide entirely with the periods I use for the model.
For the NFHS-4, I use the model results for the 2005--2016 period, 
for NFHS-3 the 1995-2004, NFHS-2 1985-1994, and for NFHS-1 1972-1984.

The predicted fertility using the hazard model assumes that women in the sample
follows the parity progression probability based on the given period. 
For example, for the NFHS-4 the predictions assume that the women in the sample follow
the estimated behavior for all four spells based on the 2005--2016 sample.

The predicted cohort fertility based on the hazard model is consistently higher and
declines faster than the predicted fertility based on the four-parity fertility rate.

The faster decline in predicted fertility based on the hazard model is consistent with
a large bias in the fertility rate measure arising from later marriage, increased use of 
sex selection, and the general increase in birth spacing shown above.

Consistent with the rapid increase in the use of sex selection from the mid-eighties
the most substantial differences between the two fertility measures are for higher
educated women for earlier samples.
